---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r codechunk1}
setwd("C:/Users/pedew_000/RepData_PeerAssessment1/")
Activity_Monitoring <- read.csv(unz("activity.zip","activity.csv"), header = TRUE, sep =",",
                                colClasses = c("numeric", "character", "numeric"))
head(Activity_Monitoring)
names(Activity_Monitoring)
```
## What is mean total number of steps taken per day?
```{r codechunk2}
Activity_Monitoring$date <- as.Date(Activity_Monitoring$date, "%Y-%m-%d")

TotalSteps <- aggregate(steps ~ date, data = Activity_Monitoring, sum, na.rm = TRUE)
head(TotalSteps)

hist(TotalSteps$steps, 
     main = "Total number of steps taken per day",
     xlab = "day", 
     col = "blue")

mean_Total_Steps <- mean(TotalSteps$steps, na.rm=TRUE)
median_Total_Steps <- median(TotalSteps$steps, na.rm=TRUE)
```

The mean total number of steps taken per day is `r mean_Total_Steps` .
The median of the total number of steps taken per day is `r median_Total_Steps`r .

## What is the average daily activity pattern?

The interval variable is the 5 minute interval identifier. (x-axis).
The average numbers of steps taken across all days is the mean of the steps taken (y-axis).

```{r codechunk3}
Time_Series_Average_Steps <- tapply(Activity_Monitoring$steps, Activity_Monitoring$interval, mean, na.rm = TRUE)

plot(row.names(Time_Series_Average_Steps),
     Time_Series_Average_Steps,
     type = "l",
     xlab = "Interval",
     ylab = "Average steps across all Days", main = "Average daily activity pattern",
     col = "blue")

Max_Interval <- which.max(Time_Series_Average_Steps)
```

The 5-minute interval that contains the maximum of steps, on average across all days, is `r names(Max_Interval)`.

## Imputing missing values
```{r codechunk4}
Missing_Values <- sum(is.na(Activity_Monitoring))

Missing_Value_Positions <- which(is.na(Activity_Monitoring$steps))

AverageSteps <- aggregate(steps ~ interval, data = Activity_Monitoring, FUN = mean)
Missing_Value_Fill <- numeric()
for (i in 1:nrow(Activity_Monitoring)) {
    Measurement <- Activity_Monitoring[i, ]
    if (is.na(Measurement$steps)) {
        steps <- subset(AverageSteps, interval == Measurement$interval)$steps
    } else {
        steps <- Measurement$steps
    }
    Missing_Value_Fill <- c(Missing_Value_Fill, steps)
}

New_Activity_Monitoring <- Activity_Monitoring
New_Activity_Monitoring$steps <- Missing_Value_Fill

New_Total_Steps <- aggregate(steps ~ date, data = New_Activity_Monitoring, sum, na.rm = TRUE)

hist(New_Total_Steps$steps, 
     main = "Total number of steps taken per day",
     xlab = "day", 
     col = "blue")
```

The total number of missing values is `r Missing_Values`.

The mean total number of steps taken per day is `r mean(New_Total_Steps$steps)` .
The median of the total number of steps taken per day is `r median(New_Total_Steps$steps)`.

## Are there differences in activity patterns between weekdays and weekends?
```{r codechunk5}
Day <- weekdays(New_Activity_Monitoring$date)
DayLevel <- vector()
for (i in 1:nrow(New_Activity_Monitoring)) {
    if (Day[i] == "Saturday") {
        DayLevel[i] <- "Weekend"
    } else if (Day[i] == "Sunday") {
        DayLevel[i] <- "Weekend"
    } else {
        DayLevel[i] <- "Weekday"
    }
}
New_Activity_Monitoring$DayLevel <- DayLevel
New_Activity_Monitoring$DayLevel <- factor(New_Activity_Monitoring$DayLevel)

StepsByDay <- aggregate(steps ~ interval + DayLevel, data = New_Activity_Monitoring, mean)
names(StepsByDay) <- c("interval", "daylevel", "steps")

library(lattice)

xyplot(steps ~ interval | daylevel, StepsByDay, type = "l", layout = c(1, 2), 
    xlab = "Interval", ylab = "Number of steps")

```